---
title: "RI-CLPM P4 - SES"
author: "Ayla Pollmann"
date: "2023-07-28"
output: 
  html_document:
    toc: true
    toc_depth: 3
---

# Introduction

In this script, we are analyzing SES (measured as income-to-needs ratio) and their relationship with mental health & brain development.

***

```{r setup, include=FALSE}
packages <- c(
  "lavaan", "sjmisc", "corrplot", "semTools", "tidyverse", "ggplot2", 
  "readxl", "semPlot", "tidyr", "dplyr", "extrafont", "reshape2", 
  "readr", "factoextra", "NbClust", "mice", "naniar", "writexl", 
  "mdatools", "PerformanceAnalytics", "jmv", "caret", "MASS", 
  "gplots", "mclust", "car", "reshape", "nnet", "kableExtra", 
  "ggeffects", "marginaleffects", "FSA", "broom", "sjlabelled", 
  "vtable", "pls", "cowplot", "patchwork", "hrbrthemes","patchwork",
  "officer","flextable"
)

installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages])
}

invisible(lapply(packages, library, character.only = TRUE))

rm(installed_packages,packages)

set.seed(245)
```

```{r, include = FALSE}
#LOAD ABCD DATA
```

# Elbow & Silhouette Method to determine number of clusters
```{r}
jpeg(file="1. Elb_Optimal Nr of Clusters.jpeg")
Elb <- fviz_nbclust(DataImp[,22:23], kmeans, method = "wss") +  #within cluster sums of squares
     labs(subtitle = "1. Elbow method")
Elb
dev.off()

jpeg(file="1. Sil_Optimal Nr of Clusters.jpeg")
Sil <- fviz_nbclust(DataImp[,22:23], kmeans, method = "silhouette")+   #average silhouette
      labs(subtitle = "1. Silhouette method")
Sil
dev.off()

rm(Elb,Sil)
```

# K-means Clustering - Trajectory & Height
```{r}
set.seed(345)
Kmeans<- kmeans(DataImp[,22:23], 2, iter.max = 10, nstart = 50) 
Kmeans$size
Kmeans$centers
summary(Kmeans)
```

```{r}
Groups <- as.factor(Kmeans$cluster)
samplesize <- 9495
id=factor(1:samplesize)
my_colors <- c("#00486b","#6591a3")

PlotCluster <- data.frame(c(DataImp$b1),     
                              c(DataImp$b2),
                              c(DataImp$b3),
                              as.factor(c(id,id)),         
                              as.factor(Groups))              
colnames(PlotCluster)<-c('~10 Years','~12 Years','~14 Years','ID','Cluster')  

BrainPlot<-melt(PlotCluster,by='ID')

BrainPlot <- ggplot(BrainPlot,aes(variable,value,group=ID,col=Cluster))+
  geom_point(position = "jitter", size=1,alpha=.3)+
  geom_line(position = "jitter", alpha=.1)+
  ylab('FA per participant')+
  xlab('Timepoints')+
  facet_grid(~factor(Cluster, levels=c("1","2")), 
             labeller = as_labeller(c("1" = "Group 1: High & Increasing", "2" = "Group 2: Low & Increasing")))+
  scale_color_manual(values = my_colors,
                     breaks =c("1","2"))+
  guides(col = FALSE)+
  theme_ipsum() +
  theme(axis.title = element_text(size = 12),
        axis.text = element_text(size = 12),
        strip.text = element_text(size = 12, face = "bold"), #cluster name
        axis.title.x = element_text(hjust = 0.5, size = 12, face = "bold"), #timepoints
        axis.title.y = element_text(hjust = 0.5, size = 12, face = "bold"), #FA
        axis.text.x = element_text(size = 9, angle = 10),
        plot.background = element_rect(fill = "white", color = NA),
        strip.text.x = element_text(hjust = 0.5))

BrainPlot + geom_smooth(method = "lm", formula = y ~ poly(x, 2), se = FALSE, aes(group = Cluster), color="#b93956") 

ggsave(file='1. Clusters.png', width=7, height=5, dpi=800)
```

# Regressions
```{r}
DataImp$Groups <- as.factor(Kmeans$cluster)
DataImp[,2:23] <- apply(DataImp[,2:23],2, function(x) as.numeric(as.character(x)))

# Regression of Cluster groups with income, interacting with gender
mod_SES <- glm(Groups ~ In1*SexOrthogonal, data = DataImp, family = binomial(link = "logit"))

# Regression of Cluster groups with income, interacting with gender
mod_Ed <- glm(Groups ~ Ed*SexOrthogonal, data = DataImp, family = binomial(link = "logit"))

# Regression of Cluster groups with income, interacting with gender
mod_ADI <- glm(Groups ~ ADI*SexOrthogonal, data = DataImp, family = binomial(link = "logit"))

# Combine
mod_All <- glm(Groups ~ (In1 + Ed + ADI) * SexOrthogonal, data = DataImp, family = binomial(link = "logit"))
```

```{r}
# Helmert coding
DataImp$GroupsOrthogonal <- as.factor(DataImp$Groups)
contrasts(DataImp$GroupsOrthogonal) <- contr.helmert(nlevels(DataImp$GroupsOrthogonal))

# Regression of mental health issues at T3 dependent on Cluster groups, interacting with gender
mod_MH <- lm(m3 ~ GroupsOrthogonal*SexOrthogonal, data = DataImp)
```

```{r}
extract_model_info <- function(model, model_name) {
  tidy_model <- tidy(model, conf.int = TRUE)
  glance_model <- glance(model)
  
  data.frame(
    Model = model_name,
    Term = tidy_model$term,
    Estimate = round(tidy_model$estimate, 3),
    CI_Lower = round(tidy_model$conf.low, 3),
    CI_Upper = round(tidy_model$conf.high, 3),
    P_Value = round(tidy_model$p.value, 3),
    AIC = round(glance_model$AIC, 2)
  )
}

results_SES <- extract_model_info(mod_SES, "Income-to-Need")
results_Ed <- extract_model_info(mod_Ed, "Education")
results_ADI <- extract_model_info(mod_ADI, "Area Deprivation")

all_results <- rbind(results_SES, results_Ed, results_ADI)

ft <- flextable(all_results)
ft <- autofit(ft)
ft <- add_header_lines(ft, "Logistic Regression Results for Cluster Group Membership")

doc <- read_docx()
doc <- body_add_flextable(doc, ft)
print(doc, target = "Regression Results SES.docx")

### For Mental Health ###

model_summary <- summary(mod_MH)
tidy_results <- tidy(mod_MH, conf.int = TRUE)
glance_results <- glance(mod_MH)

results_table <- data.frame(
  Term = tidy_results$term,
  Estimate = round(tidy_results$estimate, 3),
  Std_Error = round(tidy_results$std.error, 3),
  t_value = round(tidy_results$statistic, 3),
  P_value = round(tidy_results$p.value, 3),
  CI_Lower = round(tidy_results$conf.low, 3),
  CI_Upper = round(tidy_results$conf.high, 3)
)

ft <- flextable(results_table)
ft <- autofit(ft)
ft <- add_header_lines(ft, "Linear Regression Results: Behavioural/Emotional Issues")

ft <- add_footer_lines(ft, c(
  paste("R-squared:", round(glance_results$r.squared, 3)),
  paste("Adjusted R-squared:", round(glance_results$adj.r.squared, 3)),
  paste("F-statistic:", round(glance_results$statistic, 2), 
        "on", glance_results$df, "and", glance_results$df.residual, "DF"),
  paste("p-value:", format.pval(glance_results$p.value, digits = 3))
))

doc <- read_docx()
doc <- body_add_flextable(doc, ft)

print(doc, target = "Regression Results - MH.docx")
```

## Regression results
```{r}
options(scipen = 999)

sink("1. SES.txt")
Anova(mod_SES, type=3) 
summary(mod_SES)
zPT <- summary(mod_SES)$coefficients/summary(mod_SES)$standard.errors
zPT
pPT<- (1 - pnorm(abs(zPT), 0, 1)) * 2
pPT
null_deviance <- mod_SES$null.deviance
residual_deviance <- mod_SES$deviance
df_diff <- mod_SES$df.null - mod_SES$df.residual
chi_square <- null_deviance - residual_deviance
p_value <- pchisq(chi_square, df_diff, lower.tail = FALSE)
print("Overall Model χ² Test:")
print(paste("χ²(", df_diff, ") =", round(chi_square, 2), ", p =", format.pval(p_value, digits = 3)))
sink()

sink("1. Education.txt")
Anova(mod_Ed, type=3) 
summary(mod_Ed)
zPT <- summary(mod_Ed)$coefficients/summary(mod_Ed)$standard.errors
zPT
pPT <- (1 - pnorm(abs(zPT), 0, 1)) * 2
pPT
null_deviance <- mod_Ed$null.deviance
residual_deviance <- mod_Ed$deviance
df_diff <- mod_Ed$df.null - mod_Ed$df.residual
chi_square <- null_deviance - residual_deviance
p_value <- pchisq(chi_square, df_diff, lower.tail = FALSE)
print("Overall Model χ² Test:")
print(paste("χ²(", df_diff, ") =", round(chi_square, 2), ", p =", format.pval(p_value, digits = 3)))
sink()

sink("1. ADI.txt")
Anova(mod_ADI, type=3) 
summary(mod_ADI)
zPT <- summary(mod_ADI)$coefficients/summary(mod_ADI)$standard.errors
zPT
pPT<- (1 - pnorm(abs(zPT), 0, 1)) * 2
null_deviance <- mod_ADI$null.deviance
residual_deviance <- mod_ADI$deviance
df_diff <- mod_ADI$df.null - mod_ADI$df.residual
chi_square <- null_deviance - residual_deviance
p_value <- pchisq(chi_square, df_diff, lower.tail = FALSE)
print("Overall Model χ² Test:")
print(paste("χ²(", df_diff, ") =", round(chi_square, 2), ", p =", format.pval(p_value, digits = 3)))
sink()

sink("1. Combined.txt")
Anova(mod_All, type=3) 
summary(mod_All)
zPT <- summary(mod_All)$coefficients/summary(mod_All)$standard.errors
zPT
pPT<- (1 - pnorm(abs(zPT), 0, 1)) * 2
null_deviance <- mod_All$null.deviance
residual_deviance <- mod_All$deviance
df_diff <- mod_All$df.null - mod_All$df.residual
chi_square <- null_deviance - residual_deviance
p_value <- pchisq(chi_square, df_diff, lower.tail = FALSE)
print("Overall Model χ² Test:")
print(paste("χ²(", df_diff, ") =", round(chi_square, 2), ", p =", format.pval(p_value, digits = 3)))
sink()

sink("1. MH.txt")
print("ANOVA Table:")
print(anova(mod_MH))
print("Model Summary:")
print(summary(mod_MH))
coef_table <- summary(mod_MH)$coefficients
print("Coefficients Table:")
print(coef_table)
f_statistic <- summary(mod_MH)$fstatistic
df1 <- f_statistic[2]
df2 <- f_statistic[3]
p_value <- pf(f_statistic[1], df1, df2, lower.tail = FALSE)
print("Overall Model F-Test:")
print(paste("F(", df1, ",", df2, ") =", round(f_statistic[1], 2), 
            ", p =", format.pval(p_value, digits = 3)))
r_squared <- summary(mod_MH)$r.squared
adj_r_squared <- summary(mod_MH)$adj.r.squared
print("Model Fit:")
print(paste("R-squared =", round(r_squared, 3)))
print(paste("Adjusted R-squared =", round(adj_r_squared, 3)))
sink()

#estimate: The estimated value of the regression term.
#p.value: The two-sided p-value associated with the observed statistic.
#statistic: The value of a T-statistic to use in a hypothesis that the regression term is non-zero.
```

```{r}
tidy(mod_SES, conf.int = TRUE, exponentiate = TRUE) %>% 
  kable() %>% 
  kable_classic(html_font="Cambria", full_width = FALSE)

tidy(mod_Ed, conf.int = TRUE, exponentiate = TRUE) %>% 
  kable() %>% 
  kable_classic(html_font="Cambria", full_width = FALSE)

tidy(mod_ADI, conf.int = TRUE, exponentiate = TRUE) %>% 
  kable() %>% 
  kable_classic(html_font="Cambria", full_width = FALSE)

tidy(mod_All, conf.int = TRUE, exponentiate = TRUE) %>% 
  kable() %>% 
  kable_classic(html_font="Cambria", full_width = FALSE)

tidy(mod_MH, conf.int = TRUE, exponentiate = TRUE) %>% 
  kable() %>% 
  kable_classic(html_font="Cambria", full_width = FALSE)
```

```{r}
coef_data <- tidy(mod_All, conf.int = TRUE, conf.level = 0.95)

coef_data <- coef_data %>%
  filter(term != "(Intercept)") %>%
  mutate(term = case_when(
    term == "In1" ~ "Income-to-Need",
    term == "Ed" ~ "Parental Education",
    term == "ADI" ~ "Area Deprivation Index",
    term == "SexOrthogonal1" ~ "Gender",
    term == "In1:SexOrthogonal1" ~ "Income-to-Need × Gender",
    term == "Ed:SexOrthogonal1" ~ "Parental Education × Gender",
    term == "ADI:SexOrthogonal1" ~ "Area Deprivation Index × Gender",
    TRUE ~ term  
  ))

p <- ggplot(coef_data, aes(x = estimate, y = term)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  geom_pointrange(aes(xmin = conf.low, xmax = conf.high), 
                  position = position_dodge(width = 0.5),
                  color = "#6591a3") +
  theme_ipsum() +
  theme(
    axis.title.y = element_blank(),
    panel.grid.minor = element_blank(),
    strip.background = element_rect(fill = "lightgray", color = NA),
    strip.text = element_text(face = "bold"),
    axis.title.x = element_text(hjust = 0.5, face="bold")
  ) +
  labs(
    x = "Coefficient Estimate"
  )
print(p)
ggsave("Log Results.png", p, dpi = 600, width=5, height=6)
```


# Trajectory
```{r}
jpeg(file="2. Trajectory_Elb_Optimal Nr of Clusters.jpeg")
Elb <- fviz_nbclust(DataImp[,22], kmeans, method = "wss") +
     labs(subtitle = "2. Elbow method")
Elb
dev.off()

jpeg(file="2. Trajectory_Sil_Optimal Nr of Clusters.jpeg")
Sil <- fviz_nbclust(DataImp[,22], kmeans, method = "silhouette")+ 
      labs(subtitle = "2. Silhouette method")
Sil
dev.off()

rm(Elb,Sil)
```

# K-means Clustering
```{r}
set.seed(213)
Trajectory <- kmeans(DataImp[,22], 3, iter.max = 10, nstart = 50) 
Trajectory$size
Trajectory$centers
summary(Trajectory)
```

```{r}
#Reorder largest as first
Trajectory$cluster <- as.numeric(as.character(Trajectory$cluster))
temp <- Trajectory$cluster
Trajectory$cluster[temp == 1] <- 3
Trajectory$cluster[temp == 3] <- 1
#table(Trajectory$cluster)
```

```{r}
GroupsT <- as.factor(Trajectory$cluster)

my_colors <- c("#5a7302", "#8ab10e", "#c7db8f")

PlotCluster <- data.frame(c(DataImp$b1),     
                              c(DataImp$b2),
                              c(DataImp$b3),
                              as.factor(c(id,id)),         
                              as.factor(GroupsT))              
colnames(PlotCluster)<-c('~10 Years','~12 Years','~14 Years','ID','Groups')  

PlotCluster<-melt(PlotCluster,by='ID')

BrainPlot <- ggplot(PlotCluster,aes(variable,value,group=ID,col=Groups))+
  geom_point(position = "jitter", size=1,alpha=.3)+
  geom_line(position = "jitter", alpha=.1)+
  ylab('FA per participant')+
  xlab('Timepoints')+
  facet_grid(~factor(Groups, levels=c("1","2","3")), 
             labeller = as_labeller(c("1" = "Group 1: Increasing", "2" = "Group 2: Steep Increasing", "3" = "Group 3: Decreasing")))+
  scale_color_manual(values = my_colors,
                     breaks =c("1","2","3"))+
  guides(col = FALSE)+
  theme_ipsum() +
  theme(axis.title = element_text(size = 12),
        axis.text = element_text(size = 12),
        strip.text = element_text(size = 11, face = "bold"), #Groups name
        axis.title.x = element_text(hjust = 0.5, size = 12, face = "bold"), #timepoints
        axis.title.y = element_text(hjust = 0.5, size = 12, face = "bold"), #FA
        axis.text.x = element_text(size = 9, angle = 10),
        plot.background = element_rect(fill = "white"),
        strip.text.x = element_text(hjust = 0.5))

BrainPlot + geom_smooth(method = "lm", formula = y ~ poly(x, 2), se = FALSE, aes(group = Groups), color="#b93956") 

ggsave(file='2. Clusters_Trajectory.png', width=7, height=4, dpi=600)
```

```{r}
colors <- c("#5a7302", "#8ab10e", "#c7db8f")

TrajectoryPlot <- ggplot(PlotCluster, 
                    aes(x = variable, y = value, 
                        group = Groups, col = Groups, fill = Groups, linetype = Groups)) +
  geom_point(position = "jitter", size=1,alpha=.01)+
  geom_line(position = "jitter", alpha=.01)+
  stat_summary(fun = mean, geom = "line", aes(group = Groups), size = 1.5) + 
  stat_summary(fun = mean, geom = "point", size = 2) + 
  ylab('FA Trajectory') +
  xlab('Timepoints') +
  scale_color_manual(values = colors, 
                     labels = c("Increase","Steep increase","Decrease")) +
  scale_fill_manual(values = colors) +
  theme_ipsum() +
  theme(
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 12),
    axis.title.x = element_text(hjust = 0.5, size = 12, face = "bold"), 
    axis.title.y = element_text(hjust = 0.5, size = 12, face = "bold"),
    axis.text.x = element_text(size = 12),
    plot.background = element_rect(fill = "white", color = NA)
  ) +
  guides(linetype = FALSE, fill = FALSE)

TrajectoryPlot
ggsave(file='2. Trends_Trajectory.png', width=6, height=5, dpi=600)
```

# Regressions
```{r}
DataImp$GroupsT <- as.factor(Trajectory$cluster)
DataImp[,2:23] <- apply(DataImp[,2:23],2, function(x) as.numeric(as.character(x)))

# Regression of Cluster groups with income, interacting with gender
mod_SES <- multinom(GroupsT ~ In1*SexOrthogonal, data = DataImp)

# Regression of Cluster groups with income, interacting with gender
mod_Ed <- multinom(GroupsT ~ Ed*SexOrthogonal, data = DataImp)

# Regression of Cluster groups with income, interacting with gender 
mod_ADI <- multinom(GroupsT ~ ADI*SexOrthogonal, data = DataImp)

# Combine
mod_All <- multinom(GroupsT ~ (In1 + Ed + ADI) * SexOrthogonal, data = DataImp)

# Regression of mental health issues at T3 dependent on Cluster groups, interacting with gender
mod_MH <- lm(m3 ~ GroupsT*SexOrthogonal, data = DataImp)
```

```{r}
extract_model_info <- function(model, model_name) {
  tidy_model <- tidy(model, conf.int = TRUE)
  glance_model <- glance(model)
  
  data.frame(
    Model = model_name,
    Term = tidy_model$term,
    Estimate = round(tidy_model$estimate, 3),
    CI_Lower = round(tidy_model$conf.low, 3),
    CI_Upper = round(tidy_model$conf.high, 3),
    P_Value = round(tidy_model$p.value, 3),
    AIC = round(glance_model$AIC, 2)
  )
}

results_SES <- extract_model_info(mod_SES, "Income-to-Need")
results_Ed <- extract_model_info(mod_Ed, "Education")
results_ADI <- extract_model_info(mod_ADI, "Area Deprivation")

all_results <- rbind(results_SES, results_Ed, results_ADI)

ft <- flextable(all_results)
ft <- autofit(ft)
ft <- add_header_lines(ft, "Logistic Regression Results for Cluster Group Membership - Trajectory")

doc <- read_docx()
doc <- body_add_flextable(doc, ft)
print(doc, target = "Regression Results SES - Trajectory.docx")

### For Mental Health ###

model_summary <- summary(mod_MH)
tidy_results <- tidy(mod_MH, conf.int = TRUE)
glance_results <- glance(mod_MH)

results_table <- data.frame(
  Term = tidy_results$term,
  Estimate = round(tidy_results$estimate, 3),
  Std_Error = round(tidy_results$std.error, 3),
  t_value = round(tidy_results$statistic, 3),
  P_value = round(tidy_results$p.value, 3),
  CI_Lower = round(tidy_results$conf.low, 3),
  CI_Upper = round(tidy_results$conf.high, 3)
)

ft <- flextable(results_table)
ft <- autofit(ft)
ft <- add_header_lines(ft, "Linear Regression Results: Behavioural/Emotional Issues - Trajectory")

# Add model summary statistics
ft <- add_footer_lines(ft, c(
  paste("R-squared:", round(glance_results$r.squared, 3)),
  paste("Adjusted R-squared:", round(glance_results$adj.r.squared, 3)),
  paste("F-statistic:", round(glance_results$statistic, 2), 
        "on", glance_results$df, "and", glance_results$df.residual, "DF"),
  paste("p-value:", format.pval(glance_results$p.value, digits = 3))
))

doc <- read_docx()
doc <- body_add_flextable(doc, ft)

print(doc, target = "Regression Results - MH - Trajectory.docx")
```

## Regression results
```{r}
options(scipen = 999)

model_list <- list(
  SES = mod_SES,
  Education = mod_Ed,
  ADI = mod_ADI,
  Combined = mod_All
)

for (model_name in names(model_list)) {
  sink(paste0("2. Trajectory_", model_name, ".txt"))
  
  model <- model_list[[model_name]]
  
  print(Anova(model, type=3))
  print(summary(model))
  
  z <- summary(model)$coefficients / summary(model)$standard.errors
  p <- (1 - pnorm(abs(z), 0, 1)) * 2
  
  print("Z-scores:")
  print(z)
  print("P-values:")
  print(p)
  
  # Likelihood ratio test
  null_model <- multinom(GroupsT ~ 1, data = DataImp)
  lrt <- anova(null_model, model)
  
  print("Likelihood Ratio Test:")
  print(lrt)
  
  # Extract chi-square and p-value from LRT
  chi_square <- lrt$`LR stat.`[2]
  df <- lrt$`df`[2]
  p_value <- lrt$`Pr(Chi)`[2]
  
  print("Overall Model χ² Test:")
  print(paste("χ²(", df, ") =", round(chi_square, 2), ", p =", format.pval(p_value, digits = 3)))
  
  sink()
}

sink("2. Trajectory_MH.txt")
print("ANOVA Table:")
print(anova(mod_MH))
print("Model Summary:")
print(summary(mod_MH))
coef_table <- summary(mod_MH)$coefficients
print("Coefficients Table:")
print(coef_table)
f_statistic <- summary(mod_MH)$fstatistic
df1 <- f_statistic[2]
df2 <- f_statistic[3]
p_value <- pf(f_statistic[1], df1, df2, lower.tail = FALSE)
print("Overall Model F-Test:")
print(paste("F(", df1, ",", df2, ") =", round(f_statistic[1], 2), 
            ", p =", format.pval(p_value, digits = 3)))
r_squared <- summary(mod_MH)$r.squared
adj_r_squared <- summary(mod_MH)$adj.r.squared
print("Model Fit:")
print(paste("R-squared =", round(r_squared, 3)))
print(paste("Adjusted R-squared =", round(adj_r_squared, 3)))
sink()
```


```{r}
tidy(mod_SES, conf.int = TRUE, exponentiate = TRUE) %>% 
  kable() %>% 
  kable_classic(html_font="Cambria", full_width = FALSE)

tidy(mod_Ed, conf.int = TRUE, exponentiate = TRUE) %>% 
  kable() %>% 
  kable_classic(html_font="Cambria", full_width = FALSE)

tidy(mod_ADI, conf.int = TRUE, exponentiate = TRUE) %>% 
  kable() %>% 
  kable_classic(html_font="Cambria", full_width = FALSE)

tidy(mod_All, conf.int = TRUE, exponentiate = TRUE) %>% 
  kable() %>% 
  kable_classic(html_font="Cambria", full_width = FALSE)

tidy(mod_MH, conf.int = TRUE, exponentiate = TRUE) %>% 
  kable() %>% 
  kable_classic(html_font="Cambria", full_width = FALSE)
```

```{r}
coef_data <- tidy(mod_All, conf.int = TRUE, conf.level = 0.95)

coef_data <- coef_data %>%
  filter(term != "(Intercept)") %>%
  mutate(term = case_when(
    term == "In1" ~ "Income-to-Need",
    term == "Ed" ~ "Parental Education",
    term == "ADI" ~ "Area Deprivation Index",
    term == "SexOrthogonal1" ~ "Gender",
    term == "In1:SexOrthogonal1" ~ "Income-to-Need × Gender",
    term == "Ed:SexOrthogonal1" ~ "Parental Education × Gender",
    term == "ADI:SexOrthogonal1" ~ "Area Deprivation Index × Gender",
    TRUE ~ term  
  ))

group_labels <- c(
  "2" = "Group 1 vs Group 2 (steep increase)",
  "3" = "Group 1 vs Group 3 (decrease)"
)

p <- ggplot(coef_data, aes(x = estimate, y = term, color = y.level)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  geom_pointrange(aes(xmin = conf.low, xmax = conf.high), 
                  position = position_dodge(width = 0.5)) +
  scale_color_manual(values = c("2" = "#8ab10e", "3" ="#c7db8f"), name = "Group") +
  facet_wrap(~ y.level, ncol = 1, labeller = as_labeller(group_labels)) +
  theme_minimal() +
  theme(
    axis.title.y = element_blank(),
    legend.position = "bottom",
    panel.grid.minor = element_blank(),
    strip.background = element_rect(fill = "lightgray", color = NA),
    strip.text = element_text(face = "bold")
  ) +
  labs(
    x = "Coefficient Estimate",
    subtitle = "SES and group membership"
  ) +
  scale_x_continuous(limits = c(-0.3, 0.5), breaks = seq(-0.3, 0.5, 0.25))

print(p)
ggsave("Log Results - Trajectory.png", p, width=5, height=5, dpi=600)
```

# Height
```{r}
jpeg(file="3. Height_Elb_Optimal Nr of Clusters.jpeg")
Elb <- fviz_nbclust(DataImp[,23], kmeans, method = "wss") +  
     labs(subtitle = "3. Elbow method")
Elb
dev.off()

jpeg(file="3. Height_Sil_Optimal Nr of Clusters.jpeg")
Sil <- fviz_nbclust(DataImp[,23], kmeans, method = "silhouette")+ 
      labs(subtitle = "3. Silhouette method")
Sil
dev.off()
```

# K-means Clustering
```{r}
set.seed(234)
Height <- kmeans(DataImp[,23], 2, iter.max = 10, nstart = 50) 
Height$size
Height$centers
summary(Height)
```

```{r}
my_colors <- c("#00486b","#6591a3")

GroupsH <- as.factor(Height$cluster)

PlotCluster <- data.frame(c(DataImp$b1),     
                              c(DataImp$b2),
                              c(DataImp$b3),
                              as.factor(c(id,id)),         
                              as.factor(GroupsH))              
colnames(PlotCluster)<-c('~10 Years','~12 Years','~14 Years','ID','Cluster')  

PlotCluster<-melt(PlotCluster,by='ID')

BrainPlot <- ggplot(PlotCluster,aes(variable,value,group=ID,col=Cluster))+
  geom_point(position = "jitter", size=1,alpha=.3)+
  geom_line(position = "jitter", alpha=.1)+
  ylab('FA per participant')+
  xlab('Timepoints')+
  facet_grid(~factor(Cluster, levels=c("1","2")), 
             labeller = as_labeller(c("1" = "Group 1: High", "2" = "Group 2: Low")))+
  scale_color_manual(values = my_colors,
                     breaks =c("1","2"))+
  guides(col = FALSE)+
  theme_ipsum() +
  theme(axis.title = element_text(size = 12),
        axis.text = element_text(size = 12),
        strip.text = element_text(size = 12, face = "bold"), #cluster name
        axis.title.x = element_text(hjust = 0.5, size = 12, face = "bold"), #timepoints
        axis.title.y = element_text(hjust = 0.5, size = 12, face = "bold"), #FA
        axis.text.x = element_text(size = 9, angle = 10),
        plot.background = element_rect(fill = "white"),
        strip.text.x = element_text(hjust = 0.5))

BrainPlot + geom_smooth(method = "lm", formula = y ~ poly(x, 2), se = FALSE, aes(group = Cluster), color="#b93956") 

ggsave(file='3. Clusters_Height.png', width=6, height=4, dpi=600)
```

# Regressions
```{r}
DataImp$GroupsH <- as.factor(Height$cluster)
DataImp[,2:23] <- apply(DataImp[,2:23],2, function(x) as.numeric(as.character(x)))

# Regression of Cluster groups with income, interacting with gender
mod_SES <- glm(GroupsH ~ In1*SexOrthogonal, data = DataImp, family = binomial(link = "logit"))

# Regression of Cluster groups with income, interacting with gender
mod_Ed <- glm(GroupsH ~ Ed*SexOrthogonal, data = DataImp, family = binomial(link = "logit"))

# Regression of Cluster groups with income, interacting with gender
mod_ADI <- glm(GroupsH ~ ADI*SexOrthogonal, data = DataImp, family = binomial(link = "logit"))

# Combine
mod_All <- glm(GroupsH ~ (In1 + Ed + ADI) * SexOrthogonal, data = DataImp, family = binomial(link = "logit"))

# Regression of mental health issues at T3 dependent on Cluster groups, interacting with gender
mod_MH <- lm(m3 ~ GroupsH*SexOrthogonal, data = DataImp)
```

```{r}
extract_model_info <- function(model, model_name) {
  tidy_model <- tidy(model, conf.int = TRUE)
  glance_model <- glance(model)
  
  data.frame(
    Model = model_name,
    Term = tidy_model$term,
    Estimate = round(tidy_model$estimate, 3),
    CI_Lower = round(tidy_model$conf.low, 3),
    CI_Upper = round(tidy_model$conf.high, 3),
    P_Value = round(tidy_model$p.value, 3),
    AIC = round(glance_model$AIC, 2)
  )
}

results_SES <- extract_model_info(mod_SES, "Income-to-Need")
results_Ed <- extract_model_info(mod_Ed, "Education")
results_ADI <- extract_model_info(mod_ADI, "Area Deprivation")

all_results <- rbind(results_SES, results_Ed, results_ADI)

ft <- flextable(all_results)
ft <- autofit(ft)
ft <- add_header_lines(ft, "Logistic Regression Results for Cluster Group Membership - Level")

doc <- read_docx()
doc <- body_add_flextable(doc, ft)
print(doc, target = "Regression Results SES - Level.docx")

### For Mental Health ###

model_summary <- summary(mod_MH)
tidy_results <- tidy(mod_MH, conf.int = TRUE)
glance_results <- glance(mod_MH)

results_table <- data.frame(
  Term = tidy_results$term,
  Estimate = round(tidy_results$estimate, 3),
  Std_Error = round(tidy_results$std.error, 3),
  t_value = round(tidy_results$statistic, 3),
  P_value = round(tidy_results$p.value, 3),
  CI_Lower = round(tidy_results$conf.low, 3),
  CI_Upper = round(tidy_results$conf.high, 3)
)

ft <- flextable(results_table)
ft <- autofit(ft)
ft <- add_header_lines(ft, "Linear Regression Results: Behavioural/Emotional Issues - Level")

# Add model summary statistics
ft <- add_footer_lines(ft, c(
  paste("R-squared:", round(glance_results$r.squared, 3)),
  paste("Adjusted R-squared:", round(glance_results$adj.r.squared, 3)),
  paste("F-statistic:", round(glance_results$statistic, 2), 
        "on", glance_results$df, "and", glance_results$df.residual, "DF"),
  paste("p-value:", format.pval(glance_results$p.value, digits = 3))
))

doc <- read_docx()
doc <- body_add_flextable(doc, ft)

print(doc, target = "Regression Results - MH - Level.docx")
```

## Regression results
```{r}
options(scipen = 999)

sink("3. Height_SES.txt")
Anova(mod_SES, type=3) 
summary(mod_SES)
zPT <- summary(mod_SES)$coefficients/summary(mod_SES)$standard.errors
zPT
pPT<- (1 - pnorm(abs(zPT), 0, 1)) * 2
pPT
null_deviance <- mod_SES$null.deviance
residual_deviance <- mod_SES$deviance
df_diff <- mod_SES$df.null - mod_SES$df.residual
chi_square <- null_deviance - residual_deviance
p_value <- pchisq(chi_square, df_diff, lower.tail = FALSE)
print("Overall Model χ² Test:")
print(paste("χ²(", df_diff, ") =", round(chi_square, 2), ", p =", format.pval(p_value, digits = 3)))
sink()

sink("3. Height_Education.txt")
Anova(mod_Ed, type=3) 
summary(mod_Ed)
zPT <- summary(mod_Ed)$coefficients/summary(mod_Ed)$standard.errors
zPT
pPT <- (1 - pnorm(abs(zPT), 0, 1)) * 2
pPT
null_deviance <- mod_Ed$null.deviance
residual_deviance <- mod_Ed$deviance
df_diff <- mod_Ed$df.null - mod_Ed$df.residual
chi_square <- null_deviance - residual_deviance
p_value <- pchisq(chi_square, df_diff, lower.tail = FALSE)
print("Overall Model χ² Test:")
print(paste("χ²(", df_diff, ") =", round(chi_square, 2), ", p =", format.pval(p_value, digits = 3)))
sink()

sink("3. Height_ADI.txt")
Anova(mod_ADI, type=3) 
summary(mod_ADI)
zPT <- summary(mod_ADI)$coefficients/summary(mod_ADI)$standard.errors
zPT
pPT<- (1 - pnorm(abs(zPT), 0, 1)) * 2
null_deviance <- mod_ADI$null.deviance
residual_deviance <- mod_ADI$deviance
df_diff <- mod_ADI$df.null - mod_ADI$df.residual
chi_square <- null_deviance - residual_deviance
p_value <- pchisq(chi_square, df_diff, lower.tail = FALSE)
print("Overall Model χ² Test:")
print(paste("χ²(", df_diff, ") =", round(chi_square, 2), ", p =", format.pval(p_value, digits = 3)))
sink()

sink("3. Height_Combined.txt")
Anova(mod_All, type=3) 
summary(mod_All)
zPT <- summary(mod_All)$coefficients/summary(mod_All)$standard.errors
zPT
pPT<- (1 - pnorm(abs(zPT), 0, 1)) * 2
null_deviance <- mod_All$null.deviance
residual_deviance <- mod_All$deviance
df_diff <- mod_All$df.null - mod_All$df.residual
chi_square <- null_deviance - residual_deviance
p_value <- pchisq(chi_square, df_diff, lower.tail = FALSE)
print("Overall Model χ² Test:")
print(paste("χ²(", df_diff, ") =", round(chi_square, 2), ", p =", format.pval(p_value, digits = 3)))
sink()

sink("3. Height_MH.txt")
print("ANOVA Table:")
print(anova(mod_MH))
print("Model Summary:")
print(summary(mod_MH))
coef_table <- summary(mod_MH)$coefficients
print("Coefficients Table:")
print(coef_table)
f_statistic <- summary(mod_MH)$fstatistic
df1 <- f_statistic[2]
df2 <- f_statistic[3]
p_value <- pf(f_statistic[1], df1, df2, lower.tail = FALSE)
print("Overall Model F-Test:")
print(paste("F(", df1, ",", df2, ") =", round(f_statistic[1], 2), 
            ", p =", format.pval(p_value, digits = 3)))
r_squared <- summary(mod_MH)$r.squared
adj_r_squared <- summary(mod_MH)$adj.r.squared
print("Model Fit:")
print(paste("R-squared =", round(r_squared, 3)))
print(paste("Adjusted R-squared =", round(adj_r_squared, 3)))
sink()
```

```{r}
tidy(mod_SES, conf.int = TRUE, exponentiate = TRUE) %>% 
  kable() %>% 
  kable_classic(html_font="Cambria", full_width = FALSE)

tidy(mod_Ed, conf.int = TRUE, exponentiate = TRUE) %>% 
  kable() %>% 
  kable_classic(html_font="Cambria", full_width = FALSE)

tidy(mod_ADI, conf.int = TRUE, exponentiate = TRUE) %>% 
  kable() %>% 
  kable_classic(html_font="Cambria", full_width = FALSE)

tidy(mod_All, conf.int = TRUE, exponentiate = TRUE) %>% 
  kable() %>% 
  kable_classic(html_font="Cambria", full_width = FALSE)

tidy(mod_MH, conf.int = TRUE, exponentiate = TRUE) %>% 
  kable() %>% 
  kable_classic(html_font="Cambria", full_width = FALSE)
```

```{r}
coef_data <- tidy(mod_All, conf.int = TRUE, conf.level = 0.95)

coef_data <- coef_data %>%
  filter(term != "(Intercept)") %>%
  mutate(term = case_when(
    term == "In1" ~ "Income-to-Need",
    term == "Ed" ~ "Parental Education",
    term == "ADI" ~ "Area Deprivation Index",
    term == "SexOrthogonal1" ~ "Gender",
    term == "In1:SexOrthogonal1" ~ "Income-to-Need × Gender",
    term == "Ed:SexOrthogonal1" ~ "Parental Education × Gender",
    term == "ADI:SexOrthogonal1" ~ "Area Deprivation Index × Gender",
    TRUE ~ term  
  ))

p <- ggplot(coef_data, aes(x = estimate, y = term)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  geom_pointrange(aes(xmin = conf.low, xmax = conf.high), 
                  position = position_dodge(width = 0.5),
                  color = "#6591a3") +
  theme_ipsum() +
  theme(
    axis.title.y = element_blank(),
    panel.grid.minor = element_blank(),
    strip.background = element_rect(fill = "lightgray", color = NA),
    strip.text = element_text(face = "bold"),
    axis.title.x = element_text(hjust = 0.5, face="bold")
  ) +
  labs(
    x = "Coefficient Estimate",
    subtitle = "Effects of SES measures on group membership"
  )
print(p)
ggsave("Log Results - Level.png", p, dpi = 600, width = 10, height = 6)
```


```{r}
KMeans_Groups <- dplyr::select(DataImp,
                   "src_subject_id","In1","In2","In3","Ed","ADI","Sex","Groups","GroupsT","GroupsH","ADIRaw")

names(KMeans_Groups)[2] <- "In1Scaled"
names(KMeans_Groups)[3] <- "In2Scaled"
names(KMeans_Groups)[4] <- "In3Scaled"
names(KMeans_Groups)[5] <- "EdScaled"

write_xlsx(KMeans_Groups, "KMeans_Groups.xlsx")
```