---
title: "RI-CLPM P4 - SES"
author: "Ayla Pollmann"
date: "2023-07-28"
output: 
  html_document:
    toc: true
    toc_depth: 3
---

# Introduction

In this script, we are analyzing SES (measured as income-to-needs ratio) and their relationship with mental health & brain development.

## Overview of analysis

In this script, 

* we conducted an RI-CLPM for SES

***


```{r setup, include=FALSE}
packages <- c(
  "lavaan", "sjmisc", "corrplot", "semTools", "tidyverse", "ggplot2", 
  "readxl", "semPlot", "tidyr", "dplyr", "extrafont", "reshape2", 
  "readr", "factoextra", "NbClust", "mice", "naniar", "writexl", 
  "mdatools", "PerformanceAnalytics", "jmv", "caret", "MASS", 
  "gplots", "mclust", "car", "reshape", "nnet", "kableExtra", 
  "ggeffects", "marginaleffects", "FSA", "broom", "sjlabelled", 
  "vtable", "pls", "cowplot", "patchwork", "hrbrthemes", "ggcorrplot", 
  "patchwork"
)

installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages])
}

invisible(lapply(packages, library, character.only = TRUE))

rm(installed_packages, packages)

set.seed(245)
```


```{r, include = FALSE}
# LOAD ABCD Data
```

# SES (Income to needs)
## SES x brain connectivity x youth problems

```{r}
RICLPM_SES <- '
# Create between components (random intercepts)
RIx =~ 1*In1 + 1*In2 + 1*In3
RIy =~ 1*b1Scaled + 1*b2Scaled + 1*b3Scaled
RIz =~ 1*m1 + 1*m2 + 1*m3

# Create within-person centered variables
wx1 =~ 1*In1
wx2 =~ 1*In2
wx3 =~ 1*In3
wy1 =~ 1*b1Scaled
wy2 =~ 1*b2Scaled
wy3 =~ 1*b3Scaled
wz1 =~ 1*m1
wz2 =~ 1*m2
wz3 =~ 1*m3

# Estimate lagged effects between within-person centered variables
wx2 + wy2 + wz2 ~ wx1 + wy1 + wz1
wx3 + wy3 + wz3 ~ wx2 + wy2 + wz2

# Estimate covariance between within-person centered variables at first wave
wx1 ~~ wy1 # Covariance
wx1 ~~ wz1
wy1 ~~ wz1

# Estimate covariances between residuals of within-person centered variables 
# (i.e., innovations)
wx2 ~~ wy2
wx2 ~~ wz2
wy2 ~~ wz2
wx3 ~~ wy3
wx3 ~~ wz3
wy3 ~~ wz3

# Estimate variance and covariance of random intercepts
RIx ~~ RIx
RIy ~~ RIy
RIz ~~ RIz
RIx ~~ RIy
RIx ~~ RIz
RIy ~~ RIz

# Estimate (residual) variance of within-person centered variables
wx1 ~~ wx1 # Variances
wy1 ~~ wy1
wz1 ~~ wz1
wx2 ~~ wx2 # Residual variances
wy2 ~~ wy2
wz2 ~~ wz2
wx3 ~~ wx3
wy3 ~~ wy3
wz3 ~~ wz3
'

RICLPM_SES_fit <- lavaan::lavaan(RICLPM_SES,
                                 data = Data,
                                 missing = 'fiml', 
                                 meanstructure = TRUE, 
                                 int.ov.free = TRUE)

summary(RICLPM_SES_fit, standardized = TRUE)

fitMeasures(RICLPM_SES_fit, fit.measures = c("cfi","srmr", "rmsea", "pvalue", "chisq", "df","rmsea.ci.lower", "rmsea.ci.upper"))
```

# SES & Sex
## RICLPM

1 = Male; 2 = Female

```{r}
RICLPM_SES_fit_Sex <- lavaan::lavaan(RICLPM_SES,
                             data = Data,
                             meanstructure = TRUE, 
                             missing = 'fiml', 
                             int.ov.free = TRUE,
                             group = "Sex")

#summary(RICLPM_SES_fit_Sex, standardized = TRUE)
fitMeasures(RICLPM_SES_fit_Sex, fit.measures = c("cfi","srmr", "rmsea", "pvalue", "chisq", "df","rmsea.ci.lower", "rmsea.ci.upper"))
```

# Contstrained lagged-parameters for multi-groups
## Multigroup Peers

```{r}
MG_SES <- '
  RIx =~ 1*In1 + 1*In2 + 1*In3
  RIy =~ 1*b1Scaled + 1*b2Scaled + 1*b3Scaled
  RIz =~ 1*m1 + 1*m2 + 1*m3
  
  wx1 =~ 1*In1 
  wx2 =~ 1*In2
  wx3 =~ 1*In3
  wy1 =~ 1*b1Scaled
  wy2 =~ 1*b2Scaled
  wy3 =~ 1*b3Scaled
  wz1 =~ 1*m1
  wz2 =~ 1*m2
  wz3 =~ 1*m3
  
  # Estimate lagged effects between within-person centered variables (constrain   
  # autoregressive effects across groups) 
  wx2 ~ c(a1, a1)*wx1 + c(b1Scaled, b1Scaled)*wy1 + c(c1, c1)*wz1 
  wy2 ~ c(d1, d1)*wx1 + c(e1, e1)*wy1 + c(f1, f1)*wz1
  wz2 ~ c(g1, g1)*wx1 + c(h1, h1)*wy1 + c(i1, i1)*wz1

  wx3 ~ c(a2, a2)*wx2 + c(b2Scaled, b2Scaled)*wy2 + c(c2, c2)*wz2
  wy3 ~ c(d2, d2)*wx2 + c(e2, e2)*wy2 + c(f2, f2)*wz2
  wz3 ~ c(g2, g2)*wx2 + c(h2, h2)*wy2 + c(i2, i2)*wz2  

  wx1 ~~ wy1
  wx1 ~~ wz1
  wy1 ~~ wz1
  
  wx2 ~~ wy2 
  wx2 ~~ wz2
  wy2 ~~ wz2
  wx3 ~~ wy3
  wx3 ~~ wz3
  wy3 ~~ wz3
  
  RIx ~~ RIx
  RIy ~~ RIy
  RIz ~~ RIz
  RIx ~~ RIy
  RIx ~~ RIz
  RIy ~~ RIz

  wx1 ~~ wx1
  wy1 ~~ wy1
  wz1 ~~ wz1
  wx2 ~~ wx2
  wy2 ~~ wy2
  wz2 ~~ wz2
  wx3 ~~ wx3
  wy3 ~~ wy3
  wz3 ~~ wz3'
  

MG_Sex_fit <- lavaan(MG_SES, 
                     data = Data,
                     meanstructure = TRUE, 
                     missing = 'fiml', 
                     int.ov.free = TRUE,
                     group = "Sex")

ANOVASex <- anova(RICLPM_SES_fit_Sex, MG_Sex_fit)

write.csv(ANOVASex, file='Global_ANOVA_SES_Sex.csv', row.names = FALSE)
```

# Correlations

```{r}
CorMerged <- Data[, c("In1", "In2", "In3", "b1Scaled", "b2Scaled", "b3Scaled", "m1", "m2", "m3")]

CorMerged <- CorMerged %>%
  mutate(across(everything(), ~as.numeric(as.character(.))))

corr <- cor(CorMerged, use = "pairwise.complete.obs", method = "pearson") %>%
  round(2)

corr[is.nan(corr)] <- NA

p.mat <- cor_pmat(CorMerged, method = "pearson")

HeatmapPlot <- ggcorrplot(
  corr,
  type = "lower",
  outline.color = "white",
  lab = TRUE,
  lab_size = 3, 
  tl.cex = 10,  
  tl.srt = 45,  
  p.mat = p.mat,
  sig.level = 0.001, 
  insig = "pch",
  ggtheme = ggplot2::theme_gray,
  colors = c("#6B9AC4", "white", "#a84d5d"),
  show.legend = TRUE,
  legend.title = "Correlation"
) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

custom_labels <- c("Income T1", "Income T2", "Income T3", 
                   "WM T1", "WM T2", "WM T3", 
                   "E&B T1", "E&B T2", "E&B T3")

HeatmapPlot <- HeatmapPlot +
  scale_x_discrete(labels = custom_labels, limits = names(CorMerged)) +
  scale_y_discrete(labels = custom_labels, limits = names(CorMerged))

print(HeatmapPlot)

print(p.mat)

ggsave("Overall_Correlation_Plot.png", plot = HeatmapPlot, width = 8, height = 6, units = "in")

rm(p.mat,corr)
```

```{r}
data_long <- Data %>%
  select(src_subject_id, In1, In2, In3, m1, m2, m3) %>%
  pivot_longer(
    cols = -src_subject_id,
    names_to = c(".value", "time"),
    names_pattern = "(\\w+)(\\d)"
  ) %>%
  mutate(time = as.numeric(time))

plot_simpsons_paradox <- function(data, var1, var2) {
  overall_trend <- lm(as.formula(paste(var2, "~", var1)), data = data)
  
  ggplot(data, aes_string(x = var1, y = var2, group = "src_subject_id")) +
    geom_point(aes(color = as.factor(src_subject_id)), alpha = 0.3) +
    geom_line(aes(color = as.factor(src_subject_id)), alpha = 0.1) +
    geom_smooth(method = "lm", aes(group = 1), se = FALSE, color = "black", size = 1.5) + 
    facet_wrap(~time, ncol = 3, scales = "free") +
    theme_minimal() +
    theme(legend.position = "none") +
    labs(
      x = "Income-to-Need Ratio",
      y = "Mental Health Score",
      title = "Within-Person Trend per Timepoint (1-3)",
      subtitle = paste("Overall trend:", round(coef(overall_trend)[2], 3))
    )
}

p1 <- plot_simpsons_paradox(data_long, "In", "m")

subject_averages <- data_long %>%
  group_by(src_subject_id) %>%
  summarize(
    avg_In = mean(In, na.rm = TRUE),
    avg_m = mean(m, na.rm = TRUE)
  )

p2 <- ggplot(subject_averages, aes(x = avg_In, y = avg_m)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  theme_minimal() +
  labs(
    x = "Average Income-to-Need Ratio",
    y = "Average Mental Health Score",
    title = "Between-Person Trend",
    subtitle = "Each point represents a subject's average"
  )

combined_plot <- p1 / p2 +
  plot_layout(heights = c(2, 1)) +
  plot_annotation(
    title = "Simpsons Paradox: Income-to-Need Ratio vs Mental Health",
    subtitle = "Within-person trends and between-person averages",
    theme = theme(plot.title = element_text(size = 16, face = "bold"),
                  plot.subtitle = element_text(size = 12, face = "italic"))
  )

print(combined_plot)

ggsave("simpsons_paradox_plot.png", combined_plot, width = 8, height = 7, dpi = 500)
```

```{r}
plot_interaction <- function(data, var1, var2, title_text, x_label, y_label) {
  overall_trend <- lm(as.formula(paste(var2, "~", var1)), data = data)
  
  ggplot(data, aes_string(x = var1, y = var2, group = "src_subject_id")) +
    geom_point(aes(color = as.factor(src_subject_id)), alpha = 0.3) +
    geom_line(aes(color = as.factor(src_subject_id)), alpha = 0.1) +
    geom_smooth(method = "lm", aes(group = 1), se = FALSE, color = "black", size = 1.5) + 
    facet_wrap(~time, ncol = 3, scales = "free") +
    theme_minimal() +
    theme(legend.position = "none") +
    labs(
      x = x_label,
      y = y_label,
      title = title_text,
      subtitle = paste("Overall trend:", round(coef(overall_trend)[2], 3))
    )
}

data_In1_m2 <- data_long %>% filter(time == 1) %>% select(src_subject_id, In1 = In, m2 = m, time)
data_In2_m3 <- data_long %>% filter(time == 2) %>% select(src_subject_id, In2 = In, m3 = m, time)
data_m1_In2 <- data_long %>% filter(time == 1) %>% select(src_subject_id, m1 = m, In2 = In, time)
data_m2_In3 <- data_long %>% filter(time == 2) %>% select(src_subject_id, m2 = m, In3 = In, time)

p_In1_m2 <- plot_interaction(data_In1_m2, "In1", "m2",
                             title_text = "A: Income T1 vs Behavioural & Emotional Issues T2",
                             x_label = "Income-to-Need Ratio T1",
                             y_label = "Behavioural & Emotional Issues T2")

p_In2_m3 <- plot_interaction(data_In2_m3, "In2", "m3",
                             title_text = "B: Income T2 vs Behavioural & Emotional Issues T3",
                             x_label = "Income-to-Need Ratio T2",
                             y_label = "Behavioural & Emotional Issues T3")

p_m1_In2 <- plot_interaction(data_m1_In2, "In2", "m1",
                             title_text = "C: Behavioural & Emotional Issues T1 vs Income T2",
                             y_label = "Behavioural & Emotional Issues T1",
                             x_label = "Income-to-Need Ratio T2")

p_m2_In3 <- plot_interaction(data_m2_In3, "In3","m2",
                             title_text = "D: Behavioural & Emotional Issues T2 vs  Income T3",
                             y_label = "Behavioural & Emotional Issues T2",
                             x_label = "Income-to-Need Ratio T3")

# Combine the plots into a single figure
combined_plot <- (p_In1_m2 / p_In2_m3) | (p_m1_In2 / p_m2_In3) +
  plot_layout(guides = 'collect') +
  plot_annotation(
    title = "Interactions of Income-to-Need Ratio and Mental Health Scores",
    subtitle = "Exploring specific interactions between variables across time points",
    theme = theme(plot.title = element_text(size = 16, face = "bold"),
                  plot.subtitle = element_text(size = 12, face = "italic"))
  )

print(combined_plot)

ggsave("interaction_plots.png", combined_plot, width = 10, height = 8, dpi = 500)
```
